Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-09-13 arxiv/diff_cov.tex
revise

\begin{definition}[Limiting variance]
	Define the following
	\[ \Pi_{\vari} \define \frac{1}{n_1 + n_2}\cdot \bigtr{ (a_1 \Sigma_1 + a_2\Sigma_2)^{-1} \Sigma}. \]

@@

\begin{definition}[Limiting bias]
Recall that $a_1$ and $a_2$ are the solutions of the equation system \eqref{eq_a12extra}. Let
	\[ \Pi \define  M \frac{a_3 M^{\top}M + (1 + a_4) \id_{p\times p}}{(a_1 M^{\top}M + a_2\id_{p\times p})^2} M^{\top}, \]
where $a_{3}$ and $a_4$ are the solutions of the following two equations:
\begin{gather}\label{eq_a34extra}
		\left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_4 - b_1 \cdot  a_3
		= b_0, \quad \left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_3 -  b_1 \cdot  a_4 = b_1 .
%		\left(\frac{n_1}{\hat a_1^2} -  \sum_{i=1}^p \frac{\hat \lambda_i^4   }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_4 -\left(\sum_{i=1}^p \frac{\hat \lambda_i^2  }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_3
%		= \sum_{i=1}^p \frac{\hat \lambda_i^2 }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }. \label{eq_a4}
	\end{gather}
In the above equation, we denote $b_k$ as $\frac1{p}\sum_{i=1}^p {\lambda_i^{2k}} / { (\lambda_i^2 a_1 + a_2)^2  }$, for $k = 0, 1, 2$.
\end{definition}

@@

\begin{lemma}[Variance asymptotics]\label{lem_cov_shift}
	%Let $X_i\in\real^{n_i\times p}$ be a random matrix that contains i.i.d. row vectors with mean $0$ and variance $\Sigma_i\in\real^{p\times p}$, for $i = 1, 2$.
	%Suppose $X_1=Z_1\Sigma_1^{1/2}\in \R^{n_1\times p}$ and $X_2=Z_2\Sigma_2^{1/2}\in \R^{n_2\times p}$ satisfy Assumption \ref{assm_secA1} with $\rho_1:=n_1/p>1$ and $\rho_2:=n_2/p>1$ being fixed constants.
	%Denote by $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$ and
	Let $\Sigma \in \real^{p \times p}$ be any fixed and deterministic matrix.
	In the setting of Theorem \ref{thm_main_informal},
	with probability $1-\oo(1)$ over the randomness of $X_1$ and $X_2$, we have that %for any constant $\e>0$,
	%When $n_1 = c_1 p$ and $n_2 = c_2 p$, we have that with high probability over the randomness of $X_1$ and $X_2$, the following equation holds
	\begin{align}\label{lem_cov_shift_eq}
		\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} \Sigma} = \Pi_{\vari} + o(\norm{\Sigma}). %\bigo{\|\| p^{-1/2+\epsilon}}
	\end{align}

@@

Lemma \ref{lem_cov_shift} also implies the asymptotic limit of the prediction loss of single-task learning.
%The next lemma, which is , helps to determine the asymptotic limit of $\te(\hat{\beta}_t^{\STL})=\sigma^2   \bigtr{(X_2^{\top}X_2)^{-1}\Sigma_2}$ as $p\to \infty$.
In particular, by setting $X_1 = 0$, we obtain the following corollary, which is a well-known result in random matrix theory.
\begin{corollary}[See e.g. Theorem 2.4 in \citet{isotropic}]\label{lem_minv}
	In the setting of Lemma \ref{lem_cov_shift}, with high probability we have that
	\be\label{XXA}  \bigtr{(X_2^{\top}X_2)^{-1}A} = \frac{1}{n_2 - p} \cdot \bigtr{\Sigma_2^{-1}A} + o(\norm{A}). \ee
\end{corollary} %\bigo{ \|A\|p^{-1/2+\epsilon}}
To see this, note that when $n_1=0$, $a_1 = 0$ and $a_2 = (n_2-p) / n_2$ is the solution of \eqref{eq_a12extra}, and one can see that equation \eqref{lem_cov_shift_eq} reduces to equation \eqref{XXA}.
We briefly describe the history behind the above well-known result.
When the entries of $X_2$ are multivariate Gaussian, this result recovers the classical result for the mean of inverse Wishart distribution \cite{anderson1958introduction}.
For general non-Gaussian random matrices, it can be obtained using Stieltjes transform method; see e.g., Lemma 3.11 of \cite{bai2009spectral}.
Here we have stated a result from Theorem 2.4 in \cite{isotropic}, which gives an almost sharp concentration error bound.
One can see that our result extends Lemma \ref{lem_minv} from a single sample covariance matrix to the sum of two independent sample covariance matrices.

@@

Our next result shows that the bias equation \eqref{eq_bias_2task} is described by the above limiting bias asymptotically.
\begin{lemma}[Bias asymptotics]\label{lem_cov_derivative}
In the setting of Theorem \ref{thm_main_informal}, let $w \in \R^p$ be any vector that is independent of $X_1$ and $X_2$.
With high probability, we have that
\begin{equation}\label{lem_cov_derv_eq}
\begin{split}
\bignorm{\Sigma_2^{1/2} \bigbrace{\frac{X_1^{\top}X_1 + X_2^{\top}X_2}{n_1 + n_2}}^{-1} \Sigma_1^{1/2} w}^2
= w^{\top} \Pi w + o(\|w\|^2),
\end{split}
\end{equation}

>>>>

\begin{definition}[Limiting variance]
	Define the following
	\[ \Pi_{\vari} \define \frac{1}{n_1 + n_2}\cdot \bigtr{ (a_1 \Sigma_1 + a_2\Sigma_2)^{-1} \Sigma_2}. \]

@@

\begin{definition}[Limiting bias]
Recall that $a_1$ and $a_2$ are the solutions of the equation system \eqref{eq_a12extra}. Let
	\[ \Pi_\bias \define  M \frac{a_3 M^{\top}M + (1 + a_4) \id_{p\times p}}{(a_1 M^{\top}M + a_2\id_{p\times p})^2} M^{\top}, \]
where $a_{3}$ and $a_4$ are the solutions of the following two equations:
\begin{gather}\label{eq_a34extra}
		\left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_4 - b_1 \cdot  a_3
		= b_0, \quad \left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_3 -  b_1 \cdot  a_4 = b_1 .
%		\left(\frac{n_1}{\hat a_1^2} -  \sum_{i=1}^p \frac{\hat \lambda_i^4   }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_4 -\left(\sum_{i=1}^p \frac{\hat \lambda_i^2  }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_3
%		= \sum_{i=1}^p \frac{\hat \lambda_i^2 }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }. \label{eq_a4}
	\end{gather}
In the above equation, we denote $b_k$ as $\frac1{p}\sum_{i=1}^p {\lambda_i^{2k}} / { (\lambda_i^2 a_1 + a_2)^2  }$, for $k = 0, 1, 2$.
\end{definition}

@@

\begin{lemma}[Variance asymptotics]\label{lem_cov_shift}
	%Let $X_i\in\real^{n_i\times p}$ be a random matrix that contains i.i.d. row vectors with mean $0$ and variance $\Sigma_i\in\real^{p\times p}$, for $i = 1, 2$.
	%Suppose $X_1=Z_1\Sigma_1^{1/2}\in \R^{n_1\times p}$ and $X_2=Z_2\Sigma_2^{1/2}\in \R^{n_2\times p}$ satisfy Assumption \ref{assm_secA1} with $\rho_1:=n_1/p>1$ and $\rho_2:=n_2/p>1$ being fixed constants.
	%Denote by $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$ and
	Let $\Sigma \in \real^{p \times p}$ be any fixed and deterministic matrix.
	In the setting of Theorem \ref{thm_main_informal},
	with probability $1-\oo(1)$ over the randomness of $X_1$ and $X_2$, we have that %for any constant $\e>0$,
	%When $n_1 = c_1 p$ and $n_2 = c_2 p$, we have that with high probability over the randomness of $X_1$ and $X_2$, the following equation holds
	\begin{align}\label{lem_cov_shift_eq}
		\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} \Sigma} = \bigtr{ (a_1 \Sigma_1 + a_2\Sigma_2)^{-1} \Sigma} + o(\norm{\Sigma}). %\bigo{\|\| p^{-1/2+\epsilon}}
	\end{align}

@@

Lemma \ref{lem_cov_shift} also implies the asymptotic limit of the prediction loss of single-task learning.
%The next lemma, which is , helps to determine the asymptotic limit of $\te(\hat{\beta}_t^{\STL})=\sigma^2   \bigtr{(X_2^{\top}X_2)^{-1}\Sigma_2}$ as $p\to \infty$.
In particular, by setting $X_1 = 0$, we obtain the following corollary, which is a well-known result in random matrix theory.
\begin{corollary}[See e.g. Theorem 2.4 in \citet{isotropic}]\label{lem_minv}
	In the setting of Lemma \ref{lem_cov_shift}, with high probability we have that
	\be\label{XXA}  \bigtr{(X_2^{\top}X_2)^{-1}\Sigma} = \frac{1}{n_2 - p} \cdot \bigtr{\Sigma_2^{-1}\Sigma} + o(\norm{\Sigma}). \ee
\end{corollary} %\bigo{ \|A\|p^{-1/2+\epsilon}}
To see this, note that when $n_1=0$, $a_1 = 0$ and $a_2 = (n_2-p) / n_2$ is the solution of \eqref{eq_a12extra}, and one can see that equation \eqref{lem_cov_shift_eq} reduces to equation \eqref{XXA}.
We briefly describe the history behind the above well-known result.
When the entries of $X_2$ are multivariate Gaussian, this result recovers the classical result for the mean of inverse Wishart distribution \cite{anderson1958introduction}.
For general non-Gaussian random matrices, it can be obtained using Stieltjes transform method; see e.g., Lemma 3.11 of \cite{bai2009spectral}.
Here we have stated a result from Theorem 2.4 in \cite{isotropic}, which gives an almost sharp concentration error bound.
One can see that our result extends Lemma \ref{lem_minv} from a single sample covariance matrix to the sum of two independent sample covariance matrices.

@@

Our next result shows that the bias equation \eqref{eq_bias_2task} is described by the above limiting bias asymptotically.
\begin{lemma}[Bias asymptotics]\label{lem_cov_derivative}
In the setting of Theorem \ref{thm_main_informal}, let $w \in \R^p$ be any vector that is independent of $X_1$ and $X_2$.
With high probability, we have that
\begin{equation}\label{lem_cov_derv_eq}
\begin{split}
\bignorm{\Sigma_2^{1/2} \bigbrace{\frac{X_1^{\top}X_1 + X_2^{\top}X_2}{n_1 + n_2}}^{-1} \Sigma_1^{1/2} w}^2
= w^{\top} \Pi_\bias w + o(\|w\|^2),
\end{split}
\end{equation}
